---
title: "integrate data from adult mLN and iLN"
author: "A.DeMartin"
date: "2025-05-06"
output: 
  html_document:
    keep_md: true
    toc: true
editor_options: 
  chunk_output_type: inline
---

```{r setup, include=FALSE}
options(width = 100)
knitr::opts_chunk$set(warning = FALSE, message = FALSE, dev = c("png", "pdf"))
seed <- 1234
```

## load and preprocess
### load packages
```{r load packages, warning=FALSE, include=FALSE}
library(ExploreSCdataSeurat3)
library(runSeurat3)
library(Seurat)
library(ggpubr)
library(pheatmap)
library(SingleCellExperiment)
library(dplyr)
library(tidyverse)
library(viridis)
library(muscat)
library(circlize)
library(destiny)
library(scater)
library(metap)
library(multtest)
library(clusterProfiler)
library(org.Hs.eg.db)
library(msigdbr)
library(enrichplot)
library(DOSE)
library(grid)
library(gridExtra)
library(ggupset)
library(VennDiagram)
library(NCmisc)
```

### load object
```{r load merged object all, eval=FALSE, include=FALSE}
fileNam <- "/Users/immbio/Desktop/Project/Angelina/LNmLToRev/data/LNmLToRev_allmerged_seurat.rds"
seuratM <- readRDS(fileNam)
table(seuratM$dataset)
table(seuratM$RNA_snn_res.0.25)
table(seuratM$orig.ident)
```

### subset adult
```{r subset adult datasets, eval=FALSE, include=FALSE}
seuratA <- subset(seuratM, timepoint == "8w")
table(seuratM$timepoint)
table(seuratA$timepoint)

#rerun seurat
seuratA <- NormalizeData (object = seuratA)
seuratA <- FindVariableFeatures(object = seuratA)
seuratA <- ScaleData(object = seuratA, verbose = TRUE)
seuratA <- RunPCA(object=seuratA, npcs = 30, verbose = FALSE)
seuratA <- RunTSNE(object=seuratA, reduction="pca", dims = 1:20)
seuratA <- RunUMAP(object=seuratA, reduction="pca", dims = 1:20)
seuratA <- FindNeighbors(object = seuratA, reduction = "pca", dims= 1:20)

res <- c(0.25, 0.6, 0.8, 0.4)
for (i in 1:length(res)) {
  seuratA <- FindClusters(object = seuratA, resolution = res[i], random.seed = 1234)
}
```

### plot umaps
#### clustering
```{r umap clustering pre filter, eval=FALSE, include=FALSE}
DimPlot(seuratA, reduction = "umap", group.by = "RNA_snn_res.0.25" ,
        pt.size = 0.1, label = T, shuffle = T) +
  theme_bw() +
  theme(axis.text = element_blank(), axis.ticks = element_blank(), 
        panel.grid.minor = element_blank()) +
  xlab("umap1") +
  ylab("umap2")
```

#### location
```{r umap location pre filter, eval=FALSE, include=FALSE}
DimPlot(seuratA, reduction = "umap", group.by = "location" ,
        pt.size = 0.1, label = T, shuffle = T) +
  theme_bw() +
  theme(axis.text = element_blank(), axis.ticks = element_blank(), 
        panel.grid.minor = element_blank()) +
  xlab("umap1") +
  ylab("umap2")
```

### calculate cluster marker genes pre filter
```{r cluster marker genes pre filter, eval=FALSE, include=FALSE}
## calculate marker genes
Idents(seuratA) <- seuratA$RNA_snn_res.0.25
levels(seuratA)

markerGenes <- FindAllMarkers(seuratA, only.pos=T) %>% 
  dplyr::filter(p_val_adj < 0.01) 
```

### features pre filter
```{r featureplot sel genes pre filter, eval=FALSE}
selGenesViolin <- c("ENSMUSG00000020717.Pecam1", "ENSMUSG00000026395.Ptprc", "ENSMUSG00000031004.Mki67")

pList <- sapply(selGenesViolin, function(x){
  p <- FeaturePlot(seuratA, reduction = "umap", 
            features = x, 
            cols=c("lightgrey", "darkred"),
            order = T)+
  theme(legend.position="right")
  plot(p)
})
```


### filter object
```{r filter object, eval=FALSE}
## filter out Pecam1 (cluster #6) and Ptprc (cluster #9 and #10) Mki67 (#8) and pancreatic cells (#11)
table(seuratA$RNA_snn_res.0.25)
seuratF <- subset(seuratA, RNA_snn_res.0.25 %in% c("6", "8" ,"9", "10", "11"), invert = TRUE)
table(seuratF$RNA_snn_res.0.25)

seuratA <- seuratF
remove(seuratF)
```

### rerun seurat after filter
```{r rerun seurat after filter, eval=FALSE}
seuratA <- NormalizeData (object = seuratA)
seuratA <- FindVariableFeatures(object = seuratA)
seuratA <- ScaleData(object = seuratA, verbose = TRUE)
seuratA <- RunPCA(object=seuratA, npcs = 30, verbose = FALSE)
seuratA <- RunTSNE(object=seuratA, reduction="pca", dims = 1:20)
seuratA <- RunUMAP(object=seuratA, reduction="pca", dims = 1:20)
seuratA <- FindNeighbors(object = seuratA, reduction = "pca", dims= 1:20)

res <- c(0.25, 0.6, 0.8, 0.4)
for (i in 1:length(res)) {
  seuratA <- FindClusters(object = seuratA, resolution = res[i], random.seed = 1234)
}
```

### save object
```{r save merged seurat object adultonly, eval=FALSE}
### save seurat object
saveRDS(seuratA, file="/Users/immbio/Desktop/Project/Angelina/LNmLToRev/data/LNmLToRev_adultonly_seurat.rds")
```

## load object adult
```{r load merged object adultonly}
fileNam <- "/Users/immbio/Desktop/Project/Angelina/LNmLToRev/data/LNmLToRev_adultonly_seurat.rds"
seuratA <- readRDS(fileNam)
table(seuratA$dataset)
```

### set color vectors 
```{r set color vector}
colLoc <- c("#61baba", "#ba6161")
names(colLoc) <- unique(seuratA$location)

colLab <- c("#42a071", "#900C3F","#b66e8d", "#8F7700FF", "#61a4ba","#003C67FF",
            "#e3953d","#ab5711", "#714542", "#b6856e", "#FFC300")

names(colLab) <- c("FDC", "TRC", "TBRC", "IFRC", "medRC1" , "medRC2",
                   "PRC1", "PRC2", "Pi16+RC", "PRC3", "VSMC")

coltimepoint <- c("#440154FF", "#3B528BFF", "#21908CFF", "#5DC863FF")
names(coltimepoint) <- c("E18", "P7", "3w", "8w")

collocation <- c("#61baba", "#ba6161")
names(collocation) <- c("iLN", "mLN")
```

## plot umaps
### clustering
```{r umap clustering}
DimPlot(seuratA, reduction = "umap", group.by = "RNA_snn_res.0.25" ,
        pt.size = 0.1, label = T, shuffle = T) +
  theme_bw() +
  theme(axis.text = element_blank(), axis.ticks = element_blank(), 
        panel.grid.minor = element_blank()) +
  xlab("umap1") +
  ylab("umap2")
```

### location
```{r umap location}
DimPlot(seuratA, reduction = "umap", group.by = "location", cols = collocation,
        pt.size = 0.1, label = T, shuffle = T) +
  theme_bw() +
  theme(axis.text = element_blank(), axis.ticks = element_blank(), 
        panel.grid.minor = element_blank()) +
  xlab("umap1") +
  ylab("umap2")
```

### features
```{r featureplot sel genes}
selGenesViolin <- c("ENSMUSG00000022037.Clu", "ENSMUSG00000094686.Ccl21a",
                    "ENSMUSG00000074934.Grem1", "ENSMUSG00000050069.Grem2",
                    "ENSMUSG00000042436.Mfap4", "ENSMUSG00000071005.Ccl19",
                    "ENSMUSG00000016494.Cd34", "ENSMUSG00000001119.Col6a1",
                    "ENSMUSG00000020241.Col6a2","Rosa26eyfp.Rosa26eyfp", 
                    "ENSMUSG00000023078.Cxcl13", "ENSMUSG00000032135.Mcam")

pList <- sapply(selGenesViolin, function(x){
  p <- FeaturePlot(seuratA, reduction = "umap", 
            features = x, 
            cols=c("lightgrey", "darkred"),
            order = T)+
  theme(legend.position="right")
  plot(p)
})
```

## integrate data across location
```{r integrate, eval=FALSE, include=TRUE}
Idents(seuratA) <- seuratA$location

seurat.list <- SplitObject(object = seuratA, split.by = "location")
for (i in 1:length(x = seurat.list)) {
    seurat.list[[i]] <- NormalizeData(object = seurat.list[[i]],
                                      verbose = FALSE)
    seurat.list[[i]] <- FindVariableFeatures(object = seurat.list[[i]], 
        selection.method = "vst", nfeatures = 2000, verbose = FALSE)
}

seurat.anchors <- FindIntegrationAnchors(object.list = seurat.list, dims = 1:20)
seuratA.int <- IntegrateData(anchorset = seurat.anchors, dims = 1:20)
DefaultAssay(object = seuratA.int) <- "integrated"

## rerun seurat
seuratA.int <- ScaleData(object = seuratA.int, verbose = FALSE,
                         features = rownames(seuratA.int))
seuratA.int <- RunPCA(object = seuratA.int, npcs = 20, verbose = FALSE)
seuratA.int <- RunTSNE(object = seuratA.int, recuction = "pca", dims = 1:20)
seuratA.int <- RunUMAP(object = seuratA.int, recuction = "pca", dims = 1:20)

seuratA.int <- FindNeighbors(object = seuratA.int, reduction = "pca", dims = 1:20)
res <- c(0.6, 0.8, 0.4, 0.25)
for (i in 1:length(res)){
  seuratA.int <- FindClusters(object = seuratA.int, resolution = res[i],
                              random.seed = 1234)
}
```

### save int object adult
```{r save int object adultonly, eval=FALSE, include=TRUE}
saveRDS(seuratA.int, file="/Users/immbio/Desktop/Project/Angelina/LNmLToRev/data/LNmLToRev_adultonly_seurat.integrated.rds")
```

## load int object adult
```{r load int object adultonly}
fileNam <- "/Users/immbio/Desktop/Project/Angelina/LNmLToRev/data/LNmLToRev_adultonly_seurat.integrated.rds"
seuratA.int <- readRDS(fileNam)
```

```{r set default assay int object}
DefaultAssay(object = seuratA.int) <- "RNA"
seuratA.int$intCluster <- seuratA.int$integrated_snn_res.0.4
Idents(seuratA.int) <- seuratA.int$intCluster

colPal <- c("#DAF7A6", "#FFC300", "#FF5733", "#C70039", "#900C3F", "#b66e8d",
            "#61a4ba", "#6178ba", "#54a87f", "#25328a", "#b6856e",
            "#ba6161", "#20714a", "#0073C2FF", "#EFC000FF", "#868686FF", 
            "#CD534CFF","#7AA6DCFF", "#003C67FF", "#8F7700FF", "#3B3B3BFF",
            "#A73030FF", "#4A6990FF")[1:length(unique(seuratA.int$intCluster))]
names(colPal) <- unique(seuratA.int$intCluster)
```

### assign label
```{r assign label int clusters}
seuratA.int$label <- "label"
seuratA.int$label[which(seuratA.int$intCluster == "0")] <- "MedRC/IFRC"
seuratA.int$label[which(seuratA.int$intCluster == "1")] <- "actMedRC"
seuratA.int$label[which(seuratA.int$intCluster == "2")] <- "TBRC"
seuratA.int$label[which(seuratA.int$intCluster == "3")] <- "TRC"
seuratA.int$label[which(seuratA.int$intCluster == "4")] <- "MedRC"
seuratA.int$label[which(seuratA.int$intCluster == "5")] <- "PRC"
seuratA.int$label[which(seuratA.int$intCluster == "6")] <- "FDC/MRC" 
seuratA.int$label[which(seuratA.int$intCluster == "7")] <- "Pi16+RC"
seuratA.int$label[which(seuratA.int$intCluster == "8")] <- "VSMC"

table(seuratA.int$label)

colLab2 <- colLab
colLab <- c("#42a071", "#900C3F","#b66e8d", "#61a4ba", "#424671", "#003C67FF",
            "#e3953d", "#714542", "#b6856e", "#FFC300")

names(colLab) <- c("FDC/MRC", "TRC", "TBRC", "MedRC/IFRC", "MedRC" , "actMedRC",
                   "PRC", "Pi16+RC", "VSMC")
```

## Dimplot int data
### clustering
```{r umap int clustering}
DimPlot(seuratA.int, reduction = "umap",
        pt.size = 0.1, label = T, shuffle = T) +
  theme_bw() +
  theme(axis.text = element_blank(), axis.ticks = element_blank(), 
        panel.grid.minor = element_blank()) +
  xlab("umap1") +
  ylab("umap2")
```
### label
```{r umap int label}
DimPlot(seuratA.int, reduction = "umap", group.by = "label", cols = colLab)+
  theme_bw() +
  theme(axis.text = element_blank(), axis.ticks = element_blank(), 
        panel.grid.minor = element_blank()) +
  xlab("UMAP1") +
  ylab("UMAP2")

DimPlot(seuratA.int, reduction = "umap", group.by = "label", pt.size=0.5,
        cols = colLab, shuffle = T)+
  theme_void()

DimPlot(seuratA.int, reduction = "umap", group.by = "label", pt.size=0.5,
        cols = colLab, shuffle = T)+
  theme_void() + 
  theme(legend.position = "none") 
```

### label mLN iLN sep
####mLN
```{r umap int label sep mLN}
seuratA.intmLN <- subset(seuratA.int, location == "mLN")

DimPlot(seuratA.intmLN, reduction = "umap", group.by = "label", pt.size=0.5,
        cols = colLab, shuffle = T)+
  theme_void()

DimPlot(seuratA.intmLN, reduction = "umap", group.by = "label", pt.size=0.5,
        cols = colLab, shuffle = T)+
  theme_void() + 
  theme(legend.position = "none") 
```
####iLN
```{r umap int label sep iLN}
seuratA.intiLN <- subset(seuratA.int, location == "iLN")

DimPlot(seuratA.intiLN, reduction = "umap", group.by = "label", pt.size=0.5,
        cols = colLab, shuffle = T)+
  theme_void()

DimPlot(seuratA.intiLN, reduction = "umap", group.by = "label", pt.size=0.5,
        cols = colLab, shuffle = T)+
  theme_void() + 
  theme(legend.position = "none") 
```

### label split by location
```{r umap int label split by location}
DimPlot(seuratA.int, reduction = "umap", group.by = "label", cols = colLab,
        split.by = "location")+
  theme_bw() +
  theme(axis.text = element_blank(), axis.ticks = element_blank(), 
        panel.grid.minor = element_blank()) +
  xlab("UMAP1") +
  ylab("UMAP2")

DimPlot(seuratA.int, reduction = "umap", group.by = "label", pt.size=0.5,
        cols = colLab, split.by = "location", shuffle = T)+
  theme_void()
```

### location
```{r umap int location}
DimPlot(seuratA.int, reduction = "umap", group.by = "location", cols = colLoc,
        pt.size = 0.1, shuffle = T) +
  theme_bw() +
  theme(axis.text = element_blank(), axis.ticks = element_blank(), 
        panel.grid.minor = element_blank()) +
  xlab("umap1") +
  ylab("umap2")

DimPlot(seuratA.int, reduction = "umap", group.by = "location", pt.size=0.5,
        cols = colLoc, split.by = "location", shuffle = T)+
  theme_void()
```

### features int
```{r featureplot int sel genes}
selGenesViolin <- c("ENSMUSG00000022037.Clu", "ENSMUSG00000094686.Ccl21a",
                    "ENSMUSG00000074934.Grem1", "ENSMUSG00000050069.Grem2",
                    "ENSMUSG00000042436.Mfap4", "ENSMUSG00000071005.Ccl19",
                    "ENSMUSG00000016494.Cd34", "ENSMUSG00000024011.Pi16", 
                    "ENSMUSG00000001119.Col6a1", "ENSMUSG00000020241.Col6a2",
                    "Rosa26eyfp.Rosa26eyfp", "ENSMUSG00000023078.Cxcl13",
                    "ENSMUSG00000032135.Mcam", "ENSMUSG00000023034.Nr4a1")

pList <- sapply(selGenesViolin, function(x){
  p <- FeaturePlot(seuratA.int, reduction = "umap", 
            features = x, 
            cols=c("lightgrey", "darkred"),
            order = F)+
  theme(legend.position="right")
  plot(p)
})
```

## cluster characterization
### heatmap funcion
```{r heatmap funcion}
avgHeatmap <- function(seurat, selGenes, colVecIdent, colVecCond=NULL,
                       ordVec=NULL, gapVecR=NULL, gapVecC=NULL,cc=FALSE,
                       cr=FALSE, condCol=FALSE){
  
  selGenes <- selGenes$gene
  
  ## assay data
  clusterAssigned <- as.data.frame(Idents(seurat)) %>%
  dplyr::mutate(cell=rownames(.))
  colnames(clusterAssigned)[1] <- "ident"
  seuratDat <- GetAssayData(seurat)
  
  ## genes of interest
  genes <- data.frame(gene=rownames(seurat)) %>% 
    mutate(geneID=gsub("^.*\\.", "", gene)) %>% filter(geneID %in% selGenes)

  ## matrix with averaged cnts per ident
  logNormExpres <- as.data.frame(t(as.matrix(
    seuratDat[which(rownames(seuratDat) %in% genes$gene),])))
  logNormExpres <- logNormExpres %>% dplyr::mutate(cell=rownames(.)) %>%
    dplyr::left_join(.,clusterAssigned, by=c("cell")) %>%
    dplyr::select(-cell) %>% dplyr::group_by(ident) %>%
    dplyr::summarise_all(mean)
  logNormExpresMa <- logNormExpres %>% dplyr::select(-ident) %>% as.matrix()
  rownames(logNormExpresMa) <- logNormExpres$ident
  logNormExpresMa <- t(logNormExpresMa)
  rownames(logNormExpresMa) <- gsub("^.*?\\.","",rownames(logNormExpresMa))
  
  ## remove genes if they are all the same in all groups
  ind <- apply(logNormExpresMa, 1, sd) == 0
  logNormExpresMa <- logNormExpresMa[!ind,]
  genes <- genes[!ind,]

  ## color columns according to cluster
  annotation_col <- as.data.frame(gsub("(^.*?_)","",
                                       colnames(logNormExpresMa)))%>%
    dplyr::mutate(celltype=gsub("(_.*$)","",colnames(logNormExpresMa)))
  colnames(annotation_col)[1] <- "col1"
  annotation_col <- annotation_col %>%
    dplyr::mutate(cond = gsub(".*_","",col1)) %>%
    dplyr::select(cond, celltype)
  rownames(annotation_col) <- colnames(logNormExpresMa) 

  ann_colors = list(
      cond = colVecCond,
      celltype=colVecIdent)
  if(is.null(ann_colors$cond)){
    annotation_col$cond <- NULL
  }
  
  ## adjust order
  logNormExpresMa <- logNormExpresMa[selGenes,]
  if(is.null(ordVec)){
    ordVec <- levels(seurat)
  }
  logNormExpresMa <- logNormExpresMa[,ordVec]

  ## scaled row-wise
  pheatmap(logNormExpresMa, scale="row" ,treeheight_row = 0, cluster_rows = cr, 
         cluster_cols = cc,
         color = colorRampPalette(c("#2166AC", "#F7F7F7", "#B2182B"))(50),
         annotation_col = annotation_col, cellwidth=15, cellheight=10,
         annotation_colors = ann_colors, gaps_row = gapVecR, gaps_col = gapVecC)
}
```


### heatmap
```{r characterize clusters heatmap, fig.height=15, fig.width=5}
seuratA.int <- JoinLayers(seuratA.int)
Idents(seuratA.int) <- seuratA.int$intCluster

seuratAint_markers <- FindAllMarkers(seuratA.int, only.pos = T, logfc.threshold = 0.25)

### plot DE genes top 10 avg logFC
markerAll <- seuratAint_markers %>% group_by(cluster) %>% 
  mutate(geneID = gene) %>% top_n(10, avg_log2FC) %>%
  mutate(gene=gsub(".*\\.", "",  geneID)) %>% 
  filter(nchar(gene)>1)

grpCnt <- markerAll %>% group_by(cluster) %>% summarise(cnt=n())
gapR <- data.frame(cluster=unique(markerAll$cluster)) %>% 
  left_join(.,grpCnt, by="cluster") %>% mutate(cumSum=cumsum(cnt)) 
ordVec <- levels(seuratA.int)

pOut <- avgHeatmap(seurat = seuratA.int, selGenes = markerAll,
                  colVecIdent = colPal, 
                  ordVec=ordVec,
                  gapVecR=gapR$cumSum, gapVecC=NULL,cc=T,
                  cr=F, condCol=F)
```

## session info
```{r date and session info}
date()
sessionInfo()
```
