---
title: "iLNEYFPpos"
author: "A.DeMartin"
date: "2025-05-14"
output: 
  html_document:
    keep_md: true
    toc: true
editor_options: 
  chunk_output_type: inline
---

```{r setup, include=FALSE}
options(width = 100)
knitr::opts_chunk$set(warning = FALSE, message = FALSE, dev = c("png", "pdf"))
seed <- 1234
```

## load packages
```{r load packages, warning=FALSE, include=FALSE}
library(ExploreSCdataSeurat3)
library(runSeurat3)
library(Seurat)
library(ggpubr)
library(pheatmap)
library(SingleCellExperiment)
library(dplyr)
library(tidyverse)
library(viridis)
library(muscat)
library(circlize)
library(destiny)
library(scater)
library(metap)
library(multtest)
library(clusterProfiler)
library(org.Hs.eg.db)
library(msigdbr)
library(enrichplot)
library(DOSE)
library(grid)
library(gridExtra)
library(ggupset)
library(VennDiagram)
library(NCmisc)
library(slingshot)
library(RColorBrewer)
```

## load object EYFPonly and preprocess
```{r load merged object EYFPonly, eval=FALSE, include=TRUE}
fileNam <- "/Users/immbio/Desktop/Project/Angelina/LNmLToRev/data/LNmLToRev_EYFPonly_seurat.rds"
seuratEYFP <- readRDS(fileNam)
table(seuratEYFP$dataset)
table(seuratEYFP$RNA_snn_res.0.25)
table(seuratEYFP$orig.ident)
```

### subset iLN and rerun
```{r subset mLN, eval=FALSE, include=TRUE}

table(seuratEYFP$location)
seuratiLN <- subset(seuratEYFP, location == "iLN")
table(seuratiLN$orig.ident)

#rerun seurat
seuratiLN <- NormalizeData (object = seuratiLN)
seuratiLN <- FindVariableFeatures(object = seuratiLN)
seuratiLN <- ScaleData(object = seuratiLN, verbose = TRUE)
seuratiLN <- RunPCA(object=seuratiLN, npcs = 30, verbose = FALSE)
seuratiLN <- RunTSNE(object=seuratiLN, reduction="pca", dims = 1:20)
seuratiLN <- RunUMAP(object=seuratiLN, reduction="pca", dims = 1:20)
seuratiLN <- FindNeighbors(object = seuratiLN, reduction = "pca", dims= 1:20)

res <- c(0.25, 0.6, 0.8, 0.4)
for (i in 1:length(res)) {
  seuratiLN <- FindClusters(object = seuratiLN, resolution = res[i], random.seed = 1234)
}
```

### save object
```{r save object, eval=FALSE, include=TRUE}
### save seurat object
saveRDS(seuratiLN, file="/Users/immbio/Desktop/Project/Angelina/LNmLToRev/data/LNmLToRev_EYFP_iLN_seurat.rds")
```

## load object iLNEYFP
```{r load object iLN}
fileNam <- "/Users/immbio/Desktop/Project/Angelina/LNmLToRev/data/LNmLToRev_EYFP_iLN_seurat.rds"
seuratiLN <- readRDS(fileNam)
table(seuratiLN$dataset)
table(seuratiLN$RNA_snn_res.0.25)
table(seuratiLN$orig.ident)
```

### set color vectors 
```{r set color vector}
coltimepoint <- c("#440154FF", "#3B528BFF", "#21908CFF", "#5DC863FF")
names(coltimepoint) <- c("E18", "P7", "3w", "8w")

colPal <- c("#DAF7A6", "#FFC300", "#FF5733", "#C70039", "#900C3F", "#b66e8d",
            "#61a4ba", "#6178ba", "#54a87f",  "#25328a",
            "#b6856e", "#0073C2FF", "#EFC000FF", "#868686FF", "#CD534CFF", 
            "#7AA6DCFF", "#003C67FF", "#8F7700FF", "#3B3B3BFF", "#A73030FF",
            "#4A6990FF")[1:length(unique(seuratiLN$RNA_snn_res.0.4))]
names(colPal) <- unique(seuratiLN$RNA_snn_res.0.4)

collocation <- c("#61baba", "#ba6161")
names(collocation) <- c("iLN", "mLN")
```


## dimplot
### clustering
```{r dimplot clustering}
DimPlot(seuratiLN, reduction = "umap", group.by = "RNA_snn_res.0.4",
        cols = colPal, label = TRUE)+
  theme_bw() +
  theme(axis.text = element_blank(), axis.ticks = element_blank(), 
        panel.grid.minor = element_blank()) +
  xlab("UMAP1") +
  ylab("UMAP2")
```

### timepoint
```{r dimplot age}
DimPlot(seuratiLN, reduction = "umap", group.by = "timepoint",
        cols = coltimepoint)+
  theme_bw() +
  theme(axis.text = element_blank(), axis.ticks = element_blank(), 
        panel.grid.minor = element_blank()) +
  xlab("UMAP1") +
  ylab("UMAP2")
```

### calculate cluster marker genes
```{r marker RNA_snn_res.0.4, include=TRUE, eval=FALSE}
##cluster marker
Idents(seuratiLN) <- seuratiLN$RNA_snn_res.0.4
markerGenes <- FindAllMarkers(seuratiLN, only.pos=T) %>% 
  dplyr::filter(p_val_adj < 0.01)

#save table
write.table(markerGenes, 
            file= "/Users/immbio/Desktop/Project/Angelina/LNmLToRev/data/markerGenes_iLNEYFP_RNA_snn_res.0.4",
            sep="\t",
            quote=F,
            row.names=F,
            col.names=T)
```

## plot features
```{r featureplot sel genes}
selGenesViolin <- c("ENSMUSG00000022037.Clu", "ENSMUSG00000094686.Ccl21a",
                    "ENSMUSG00000074934.Grem1", "ENSMUSG00000050069.Grem2",
                    "ENSMUSG00000042436.Mfap4", "ENSMUSG00000071005.Ccl19",
                    "ENSMUSG00000016494.Cd34", "ENSMUSG00000001119.Col6a1",
                    "ENSMUSG00000020241.Col6a2","Rosa26eyfp.Rosa26eyfp", 
                    "ENSMUSG00000023078.Cxcl13", "ENSMUSG00000037405.Icam1", "ENSMUSG00000027962.Vcam1")

pList <- sapply(selGenesViolin, function(x){
  p <- FeaturePlot(seuratiLN, raster = FALSE ,reduction = "umap", 
            features = x, 
            cols=c("lightgrey", "darkred"),
            order = T)+
  theme(legend.position="right")
  plot(p)
})
```

```{r featureplot sel genes-2}
genes <- data.frame(gene=rownames(seuratiLN)) %>% 
    mutate(geneID=gsub("^.*\\.", "", gene)) 

selGenesAll <- data.frame(geneID=c("Fbln1", "Col15a1", "Cnn1", "Acta2", "Rgs5",
                                   "Cox4i2", "Pi16", "Cd34", "Emp1", "Ogn",
                                   "Fhl2", "Mcam")) %>% 
  left_join(., genes, by = "geneID") 

pList <- sapply(selGenesAll$gene, function(x){
p <- FeaturePlot(seuratiLN, reduction = "umap", 
            features = x,
            cols=c("lightgrey", "darkred"),
            order = F)+
  theme(legend.position="right")
  plot(p)
})
```

```{r featureplot sel genes pre filter, eval=FALSE}
selGenesViolin <- c("ENSMUSG00000026395.Ptprc", "ENSMUSG00000031004.Mki67", "ENSMUSG00000063011.Msln", "ENSMUSG00000045680.Tcf21")

pList <- sapply(selGenesViolin, function(x){
  p <- FeaturePlot(seuratiLN, reduction = "umap", 
            features = x, 
            cols=c("lightgrey", "darkred"),
            order = T)+
  theme(legend.position="right")
  plot(p)
})
```

## filter 
### filter object 3
```{r filter object 3, eval=FALSE, include=TRUE}
## filter out and Ptprc (cluster #10),
## and proliferating cells (cluster #7 and #8)

table(seuratiLN$RNA_snn_res.0.4)
seuratF <- subset(seuratiLN, RNA_snn_res.0.4 %in% c("10", "8", "7"), invert = TRUE)
table(seuratF$RNA_snn_res.0.4)

seuratiLNf3 <- seuratF
remove(seuratF)
table(seuratiLNf3$orig.ident)
```

### rerun after filter 3
```{r rerun mLNf3, eval=FALSE, include=TRUE}
#rerun seurat
seuratiLNf3 <- NormalizeData (object = seuratiLNf3)
seuratiLNf3 <- FindVariableFeatures(object = seuratiLNf3)
seuratiLNf3 <- ScaleData(object = seuratiLNf3, verbose = TRUE)
seuratiLNf3 <- RunPCA(object=seuratiLNf3, npcs = 30, verbose = FALSE)
seuratiLNf3 <- RunTSNE(object=seuratiLNf3, reduction="pca", dims = 1:20)
seuratiLNf3 <- RunUMAP(object=seuratiLNf3, reduction="pca", dims = 1:20)
seuratiLNf3 <- FindNeighbors(object = seuratiLNf3, reduction = "pca", dims= 1:20)

res <- c(0.25, 0.6, 0.8, 0.4)
for (i in 1:length(res)) {
  seuratiLNf3 <- FindClusters(object = seuratiLNf3, resolution = res[i], random.seed = 1234)
}
```


## load object mLNEYFPf3
```{r load object mLNf3}
fileNam <- "/Users/immbio/Desktop/Project/Angelina/LNmLToRev/data/LNmLToRev_EYFP_iLNf3_seurat.rds"
seuratiLNf3 <- readRDS(fileNam)
table(seuratiLNf3$dataset)
```

## dimplot
### clustering
```{r dimplot clustering mLNf3}
colPal <- c("#DAF7A6", "#FFC300", "#FF5733", "#C70039", "#900C3F", "#b66e8d",
            "#61a4ba", "#6178ba", "#54a87f",  "#25328a",
            "#b6856e", "#0073C2FF", "#EFC000FF", "#868686FF", "#CD534CFF", 
            "#7AA6DCFF", "#003C67FF", "#8F7700FF", "#3B3B3BFF", "#A73030FF",
            "#4A6990FF")[1:length(unique(seuratiLNf3$RNA_snn_res.0.4))]
names(colPal) <- unique(seuratiLNf3$RNA_snn_res.0.4)

DimPlot(seuratiLNf3, reduction = "umap", group.by = "RNA_snn_res.0.4",
        cols = colPal, label = TRUE)+
  theme_bw() +
  theme(axis.text = element_blank(), axis.ticks = element_blank(), 
        panel.grid.minor = element_blank()) +
  xlab("UMAP1") +
  ylab("UMAP2")
```

### timepoint
```{r dimplot timepoint mLNf3}
DimPlot(seuratiLNf3, reduction = "umap", group.by = "timepoint",
        cols = coltimepoint)+
  theme_bw() +
  theme(axis.text = element_blank(), axis.ticks = element_blank(), 
        panel.grid.minor = element_blank()) +
  xlab("UMAP1") +
  ylab("UMAP2")
```

### label transfer
```{r transfer label}
fileNam <- "/Users/immbio/Desktop/Project/Angelina/LNmLToRev/data/LNmLToRev_adultonly_seurat.integrated.rds"
seuratLab <- readRDS(fileNam)

seuratLab <- subset(seuratLab, location=="iLN")
seuratLab <- subset(seuratLab, EYFP=="pos")
table(seuratLab$label)
```


```{r transfer label-2}
labCells <- data.frame(label=seuratLab$label) %>% rownames_to_column(., "cell")
allCell <- data.frame(cell=colnames(seuratiLNf3)) %>% 
  left_join(., labCells, by= "cell")
allCell$label[which(is.na(allCell$label))] <- "unassigned"
seuratiLNf3$label <- allCell$label

table(seuratiLNf3$timepoint)
table(seuratiLNf3$label)

colLab <- c("#42a071", "#900C3F","#b66e8d", "#61a4ba", "#424671", "#003C67FF",
            "#e3953d", "#714542", "#b6856e", "#a4a4a4")

names(colLab) <- c("FDC/MRC", "TRC", "TBRC", "MedRC/IFRC", "MedRC" , "actMedRC",
                   "PRC", "Pi16+RC", "VSMC", "unassigned")

DimPlot(seuratiLNf3, reduction = "umap", group.by = "label",
        cols = colLab, shuffle=T)+
  theme_bw() +
  theme(axis.text = element_blank(), axis.ticks = element_blank(), 
        panel.grid.minor = element_blank()) +
  xlab("UMAP1") +
  ylab("UMAP2")
```

```{r transfer label-3}
DimPlot(seuratiLNf3, reduction = "umap", group.by = "label", pt.size=0.5,
        cols = colLab, order = TRUE)+
  theme_void()

DimPlot(seuratiLNf3, reduction = "umap", group.by = "label", pt.size=0.5,
        cols = colLab, shuffle = FALSE)+
  theme_void()
```

### calculate cluster marker genes iLNf3
```{r marker RNA_snn_res.0.4 iLNf3, include=TRUE, eval=FALSE}
##cluster marker
Idents(seuratiLNf3) <- seuratiLNf3$RNA_snn_res.0.4
markerGenes <- FindAllMarkers(seuratiLNf3, only.pos=T) %>% 
  dplyr::filter(p_val_adj < 0.01)

#save table
write.table(markerGenes, 
            file= "/Users/immbio/Desktop/Project/Angelina/LNmLToRev/data/markerGenes_iLNEYFPf3_RNA_snn_res.0.4",
            sep="\t",
            quote=F,
            row.names=F,
            col.names=T)
```

## plot features
### sel markers mLNf3
```{r vis markers mLNf3}
genes <- data.frame(gene=rownames(seuratiLNf3)) %>% 
    mutate(geneID=gsub("^.*\\.", "", gene)) 

selGenesAll <- data.frame(geneID=c("Rosa26eyfp","tdTomato", "Ccl19", "Ccl21a", "Cxcl13",
                                   "Fbln1", "Col15a1", "Cnn1", "Acta2","Myh11", "Rgs5",
                                   "Cox4i2", "Pi16", "Cd34", "Emp1", "Ogn","Des",
                                   "Fhl2")) %>% 
  left_join(., genes, by = "geneID") 

pList <- sapply(selGenesAll$gene, function(x){
p <- FeaturePlot(seuratiLNf3, reduction = "umap", 
            features = x,
            cols=c("lightgrey", "darkred"),
            order = F) +
  theme(legend.position="right")
  plot(p)
})
```

### save object
```{r save object mLNf3, eval=FALSE, include=TRUE}
### save seurat object
saveRDS(seuratiLNf3, file="/Users/immbio/Desktop/Project/Angelina/LNmLToRev/data/LNmLToRev_EYFP_iLNf3_seurat.rds")
```

## slingshot v2
```{r slingshot iLNf3 v2, eval=FALSE, include=TRUE}
sce <- as.SingleCellExperiment(seuratiLNf3)
sc## include endpoint in slingshot analysis
sce <- slingshot(sce, clusterLabels = 'RNA_snn_res.0.4', reducedDim = 'UMAP',
                 start.clus = "6", end.clus = c("0", "4", "7", "8"),
                 dist.method="simple", extend = 'n', stretch=0)
```

### load sce iLNf3 slingshot v2
```{r load sce iLNf3 slingshot v2}
fileNam <- "/Users/immbio/Desktop/Project/Angelina/LNmLToRev/data/LNmLToRev_EYFP_iLNf3_slingshot_v2_sce.rds"
sceiLNf3v2<- readRDS(fileNam)
```

```{r slingshot iLNf3-2 v2}
clustDat <- data.frame(clustCol=colPal) %>% rownames_to_column(., "cluster")
timepointDat <- data.frame(ageCol=coltimepoint) %>% rownames_to_column(., "timepoint")
colDat <- data.frame(cluster=sceiLNf3v2$RNA_snn_res.0.4) %>%
  mutate(timepoint=sceiLNf3v2$timepoint) %>% left_join(., clustDat, by="cluster") %>% 
  left_join(., timepointDat, by="timepoint")
```

```{r slingshot-1 iLNf3 v2}
plot(reducedDims(sceiLNf3v2)$UMAP, col = colDat$clustCol, pch=16, asp = 1)
lines(SlingshotDataSet(sceiLNf3v2), lwd=2, type = 'lineages', col = 'black')
```

```{r slingshot-2 iLNf3 v2}
plot(reducedDims(sceiLNf3v2)$UMAP, col = colDat$ageCol, pch=16, asp = 1)
lines(SlingshotDataSet(sceiLNf3v2), lwd=2, type = 'lineages', col = 'black')
```

```{r slingshot-3 iLNf3 v2}
plot(reducedDims(sceiLNf3v2)$UMAP, col = colDat$clustCol, pch=16, asp = 1)
lines(SlingshotDataSet(sceiLNf3v2), lwd=2, col='black')
```

```{r slingshot-4 iLNf3 v2}
plot(reducedDims(sceiLNf3v2)$UMAP, col = colDat$ageCol, pch=16, asp = 1)
lines(SlingshotDataSet(sceiLNf3v2), lwd=2, col='black')
```

### save slingshot object v2
```{r save slingshot iLNf3 v2, eval=FALSE, include=TRUE}
saveRDS(sce, file= "/Users/immbio/Desktop/Project/Angelina/LNmLToRev/data/LNmLToRev_EYFP_iLNf3_slingshot_v2_sce.rds")
```

```{r}
summary(sceiLNf3v2$slingPseudotime_1)
summary(sceiLNf3v2$slingPseudotime_2)
summary(sceiLNf3v2$slingPseudotime_3)
summary(sceiLNf3v2$slingPseudotime_4)
summary(sceiLNf3v2$slingPseudotime_5)
```
```{r}
colors <- colorRampPalette(rev(brewer.pal(11,'Spectral')))(100)
plotcol <- colors[cut(slingAvgPseudotime(SlingshotDataSet(sceiLNf3v2)), breaks=100)]

plot(reducedDims(sceiLNf3v2)$UMAP, col = plotcol, pch=16, asp = 1)
lines(SlingshotDataSet(sceiLNf3v2), lwd=2, col='black')
```

```{r}
colors <- colorRampPalette(brewer.pal(11,'YlOrRd'))(100)
plotcol <- colors[cut(slingAvgPseudotime(SlingshotDataSet(sceiLNf3v2)), breaks=100)]

plot(reducedDims(sceiLNf3v2)$UMAP, col = plotcol, pch=16, asp = 1)
lines(SlingshotDataSet(sceiLNf3v2), lwd=2, col='black')
```

```{r}
### color lineages
colLin <- c("#42a071","#900C3F","#e3953d", "#424671", "#b6856e")
names(colLin) <- c("1", "2", "3", "4", "5")

plot(reducedDims(sceiLNf3v2)$UMAP, col = "#bfbcbd", pch=16, asp = 1)
lines(SlingshotDataSet(sceiLNf3v2), lwd=4, col=colLin)
```

## session info
```{r date and session info}
date()
sessionInfo()
```
